\textbf{Are regime changes always bad economics? Evidence from daily financial data}

### 2019-05-10 doc/paper.tex
Wordsmithing, reformatting, and additions to main text

\singlespacing
\begin{abstract}
We examine the relationship between political instability and daily returns of national stock indices in every country that experienced a coup, assassination, or resignation for which daily data is available. A common hypothesis is that political instability discourages investment and reduces economic growth. By contrast, we find that while financial volatility increases dramatically following ``irregular'' regime changes, different types of regime change have disparate effects on stock returns. We use daily financial data and a constant mean return event study model to show that abnormal returns following resignations are large and positive (4\%), while abnormal returns following assassinations are negative and smaller in magnitude (2\%). The impact of coups tends to be negative (2\%), but some events result in positive abnormal returns of 10\% or more. We also find that volatility increases during times of protest preceding resignations, but that no clear directionality is present. We therefore find that the expected direction and magnitude of abnormal returns is dependent on the type of political event and its expected impact on economic policy.
\end{abstract}
\doublespacing

@@

In this study, we examine the impact of ``irregular'' regime changes and public protests on economic performance and remain agnostic about the direction of the effect. In contrast to the aforementioned cross country studies but like \citet{girardi2018institution} we  conduct our analysis using daily financial data. Market expectations in the form of stock market returns are used as an indicator of whether the market views different types of potentially destabilizing political events as ``good'' for economic growth. The advantage of this event study approach is that it allows us to use time-series variation in stock returns to estimate treatment effects and mitigate the endogeneity problems that are nearly ubiquitous in cross country regressions.\footnote{A growing body of work uses event studies to assess political phenomenon.  For example, political events have been used to estimate the effect of political connections on firm value \citep[e.g.][]{fisman2001estimating,faccio2006politically,goldman2009politically}. Studies in ``forensic economics" have used abnormal returns from political events to locate or examine transactions such as insider trading \citep{dube2011coups}, illegal arms trading \citep{dellavigna2010detecting}, and the impact of hostilities on the financial performance of diamond mining firms in Angola \citep{guidolin2007diamonds}.} We analyze 13 coups, 8 assassinations, 15 forced resignations, 11 public protests, and 1 natural experiment provided by the failed 2002 coup in Venezuela. Our study is therefore not only able to use an event-study approach to enhance the reliability of the estimates, but is not limited to a single case study as in \citet{girardi2018institution}. The relatively large sample enhances external validity and allows us to show that different types of political events have disparate economic effects.

We find that coups, assassinations, resignations, and public protests cause large increases in financial volatility. Abnormal returns following resignations are large and positive (4\%) while abnormal returns following assassinations are negative and smaller in magnitude (2\%). The average impact of coups is negative (2\%); however, stock prices increased dramatically during the failed overthrow of Hugo Chavez's socialist regime and decreased by approximately the same magnitude following his return to leadership. The effect of public protests on mean stock returns is uncertain, with some protests such as those during the Egpytian Revolution depressing them and others have the opposite effect.

@@

The failed coup attempt against Hugo Chavez allows us to examine the effects of both pro-business and anti-business regime change separately from simple uncertainty because investors reacted to an expected regime change twice: first, when Chavez was ousted, and second, when he was reinstated. On the evening of April 11, 2002, coup plotters removed Chavez from office and later detained him. Pedro Carmona, a Venezuelan economist and business leader, was named the transitional President of Venezuela. Two days later, on April 13, 2002, a popular uprising led to Chavez's reinstatement as president. This provides an estimate of the market's valuation of a transition from the Chavez regime to the Carmona regime and its valuation of a transition from the Carmona regime back to the Chavez regime. By extension, it provides an estimate of the impact of a shift from a left-wing populist government to a pro business regime in an emerging market.

\autoref{fig:AR-Ven} provides graphical evidence on the effect of the coup attempt. The top panel shows CARs for the 10 days prior to and following the event, along with 95\% confidence intervals. The daily ARs and corresponding confidence intervals are displayed in the bottom panel. The abnormal return on April 12, the first trading day in which investors could react to the coup, was 10\%. The market reacted similarly, albeit in the opposite direction, to Chavez's reinstatement as president: the abnormal return on the next trading day, April 15, 2002, was -8\%.

The results in the figure are particularly striking given the discrepancy between the ARs on event days 0 and 1 and all other days. The only days on which the ARs are statistically different from zero is on event days 0 and 1 after the coup attempt. The almost 0\% 10-day CAR preceding the coup makes this an ideal case as it implies that investors were unaware of the coup plot. The unexpected nature of the event means that the abnormal returns capture the true value of the regime change from Chavez to Carmona more accurately than they otherwise would.

@@

The results in \autoref{tab:AR-ass} are produced from analyses identical to those table \autoref{tab:AR-coups} but for successful assassinations rather than coups. Like the majority of coups, there is evidence that assassinations decrease stock prices. The mean event day abnormal return is negative and statistically different than zero. However, the result is driven by five events: the shooting of U.S. President William McKinley on September 6, 1901; the assassination of U.S. President John F. Kennedy on November 22, 1963; the assassination of Indian Prime Minister Indira Gandhi on October 31, 1984; the suicide bombing that killed Sri Lankan president Ranasinghe Premadasa on May 1, 1993; and the assassination of Israeli Prime Minister Yitzhak Rabin on the evening of November 4, 1995.

@@

These results are consistent with our hypothesis that the nature of the political event and its expected impact on policy matters. While the mean effect of assassinations is negative, it is smaller in magnitude than for coups. Unlike a coup, an assassination is typically an unexpected event that may not necessarily cause immediate change in economic policy. As such, we would expect CARs to be negative due to increased instability and uncertainty, but smaller in magnitude to a coup or resignation due to greater expectations of policy inertia.

@@

In contrast to coups and assassinations, abnormal returns following resignations are large and positive (see \autoref{tab:AR-resignations}). The mean event day abnormal return is over 4\% and the positive returns are persistent and grow larger over time (mean 20-day CAR $\approx$ 11\%). Furthermore, event day ARs are only negative and statistically significant at even the ten percent level in two out of the fifteen resignations (Pakistan on April 19, 1993 and Tunisia on June 31, 2011).

@@

For example, consider Ferdinand Marcos' resignation from office as President of the Philippines in February 1986. Prior to his resignation, the Philippine regime was known for rampant corruption, crony capitalism, extreme inequality, high unemployment, failed import substitution industrialization policy, and oligarchic control of the economy \citep{overholt1986rise, traywick2014}. In fact, the Philippines was the least preferred site for foreign investment amongst East Asian capitalist economies and possessed one of the worst capital investment to economic output ratios in Asia \citep{overholt1986rise}. Marcos held a snap presidential election on February 7, 1986, in which he declared victory despite overwhelming evidence of electoral fraud. Public protests ensued, and two weeks later the military withdrew its support of the Marcos regime \citep{lee2009armed}. Marcos was replaced by his electoral opponent, Corazon Aquino, who had run on a platform of economic liberalization and elimination of crony capitalism \citep{villegas1987philippines}.

@@

The short-term impact of the Egyptian Revolution on the economy was disastrous. As shown in \autoref{fig:CAR-Egypt}, abnormal returns on the Egyptian Stock Exchange Index (EGX 30) were around -7\% on January 26th and -10\% the day after. To prevent further decline during the uprising, the Egyptian Stock Exchanged closed at the end of trading on January 27th. President Mubarak resigned on February 11, but the market remained closed until March 23, when CARs declined by another 9\%, before rebounding slightly thereafter.

@@

\begin{figure}[htb]
\includegraphics{../figs/mean-volatility.pdf}
\caption{Mean of volatility estimates from GARCH(1,1) models}
\label{fig:volatility}
\end{figure}

@@

\begin{figure}
\centering
\begin{subfigure}{\textwidth}
\centering
\includegraphics[max size={.8\textwidth}{.8\textheight}]{../figs/mean-car-by-regime-change-type.pdf}
\caption{Actual event date} \label{subfig:mean-car-by-regime-change-actual}
\end{subfigure}
\begin{subfigure}{\textwidth}
\centering
\includegraphics[max size={.8\textwidth}{.8\textheight}]{../figs/mean-car-by-regime-change-type-placebo.pdf}
\caption{Event date shifted forward by one year} \label{subfig:mean-car-by-regime-change-placebo}
\end{subfigure}
\caption{Mean cumulative abnormal returns by type of regime change}\label{fig:mean-car-by-regime-change}
\end{figure}

To ensure that results of the placebo test based on 1-year did not occur by chance, \autoref{fig:mean-event-day-ar-by-regime-change} plots event-day abnormal against the number of days shifted. There are a few instances in which there are statistically significant (at the 95\% level) ARs in the same direction as those on the actual event day, but they are always smaller in magnitude than the ARs estimated using the actual event date. Most of the statistically significant placebo estimates also occur when dates are shifted within the post-event window (days 0-20), a period during which stock returns remain volatile and, in the case of resignations, there is a consistent upward trend in the CARs. When dates are shifted forward further ($\geq$ 110 days), the event day AR is only statistically significant in one case (day 365 for assassinations). Overall, these results reinforce the main results: the ARs on the actual event date capture most of effect of the regime change, although effects can sometimes persist in the short event window following the event date.

These figures also provide evidence on the extent to which regime changes appear to be unexpected. For instance, the CARs prior to assassinations and coups presented in \autoref{subfig:mean-car-by-regime-change-actual} tend to be close to zero, suggesting that investors were unaware that a negative event was likely to occur in the coming days. For resignations, CARs trend downward in the 10 days prior to the event; however, if investors were aware that a resignation were about to occur one would expect the pre-event CARs to be positive given the positive CARs observed in the post-event window. Overall, there is thus little evidence that the CARs in the post-event window are not capturing most of the effect caused by the regime changes.

@@

\begin{table}[!ht]
\caption{Non-parametric tests of the impact of regime changes} \label{tab:non-parametric}
\vspace{-5pt}
\footnotesize
\begin{center}
\begin{threeparttable}
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}d{1.3}d{1.3}d{1.3}d{1.3}d{1.3}d{1.3}d{1.3}}
  \hline
  \hline
&\multicolumn{3}{c}{Regime Change Country}&\multicolumn{3}{c}{Synthetic Control Portfolio}&\multicolumn{1}{c}{\multirow{2}{*}{Wilcoxon}}\\
\cmidrule(r){2-4} \cmidrule(r){5-7}
&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Rank}&\multicolumn{1}{c}{Sign}
&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Rank}&\multicolumn{1}{c}{Sign}
&\multicolumn{1}{c}{Rank Test}\\
\multicolumn{1}{c}{Event Type}&\multicolumn{1}{c}{CAR (0,0)}&\multicolumn{1}{c}{p-value}&\multicolumn{1}{c}{p-value}
&\multicolumn{1}{c}{CAR (0,0)}&\multicolumn{1}{c}{p-value}&\multicolumn{1}{c}{p-value}
&\multicolumn{1}{c}{p-Value}\\
  \hline
\ExpandableInput{../tables/non-parametric-ar0.txt}
   \hline
   \hline
\end{tabular*}
\scriptsize
Notes: Estimates for assassinations do not include the assassination of U.S. president William McKinley in 1901 because no control portfolios are available. Two-sided 
\end{threeparttable}
\end{center}
\end{table}

As shown in \autoref{tab:non-parametric}, the mean event day abnormal returns for coups, assassinations and resignations are all statistically different from zero at the 1\% level using the rank test statistic and the abnormal returns for coups and assassinations are significant at at least the 10\% level using the sign test. In addition, abnormal absolute returns for all events are statistically significant at the 5\% level using both the rank and sign statistics. On the other hand, the event day abnormal returns for the control portfolio are never statistically different from zero at even the 10\% level. Finally, the difference in means between the regime change country and the control portfolio are statistically different from zero for coups (1\% level), assassinations (10\% level), resignations (5\% level), and all events combined (1\% level) when using two-sided p-values from the Wilcoxon rank test.\footnote{The Wilcoxon rank test is a non-parametric statistical technique that can be used to compare differences between matched samples.} In sum, the synthetic control and small sample tests therefore suggest that the main results are not a result of deviations from normality or confounding world events.

\section{Conclusion and Discussion}
Max Weber (\citeyear{weber1947theory}) developed the concept of ``charismatic leadership'' whereby a leader's authority stems from their charisma. Such rule is often found in autocracies with weak institutions and few constrains placed on the power of leaders. \citet{jones2005leaders} has shown that leaders can have particularly large impacts on economic growth in these settings. It is therefore not surprising that there is often a strong temptation to overthrow "bad" leaders, even by undemocratic means.

@@

The results are consistent with the idea that perceptions of government competence and changes in government have large impacts on investor confidence. But although the effect of regime changes and protests on stock volatility is substantial, the effect on the direction of stock returns is less certain. Abnormal returns following resignations are large and positive (4\%), abnormal returns following assassinations are negative and smaller in magnitude (2\%), and abnormal returns following coups also tend to be negative (2\%). However, the returns surrounding the failed 2002 coup that temporarily ousted Hugo Chavez from power in Venezuela show that coups can sometimes have large positive economic effects, which is consistent with estimates from \citet{girardi2018institution} showing  the election of Allende in Chile and the subsequent coup replacing him increased stock market valuations. In total, the evidence consequently suggests that abnormal returns tend to be positive following resignations because the regime change likely replaced ``bad'' leaders, and coups can result in positive returns in settings where socialist governments are replaced with more pro-business ones. Conversely, autocratic coups may increase fear of capital appropriation or the implementation of poor economic policies, while assassinations may be merely regarded as highly uncertain events.

As in any study, there are a number potential avenues for future research. First, more research is needed to identify the pro- and anti-market characteristics of regime changes. These may be the types of regime changes used in this paper or more nuanced characteristics. Ideally, researchers would be better able to predict the expected effects of future regime changes on stock returns. Additionally, it would be useful to identify the mechanisms through which regime changes effect stock returns. In the case of Chile,  \citet{girardi2018institution} shows that increasing stock market valuations were caused by changes in private property rights rather than economic growth prospects or wage costs. Research focusing on the mechanisms in other settings would help generalize these findings. Lastly, it would be helpful to determine the extent to which stock market returns translate to broader economic development outcomes. \citet{meyersson2016political} has made progress on this front by examining the impact of coups on a number of outcomes in addition to economic growth including investment, debt, inflation, infant mortality, and years of schooling. It would be fruitful to examine whether the direction of the effects of different types of regime changes on these outcomes are consistent with their stock market effects. Such research can help enlarge the body of evidence on the extent to which regime changes cause institutional and political change, and, in turn, have significant consequences for economic development.

@@

\begin{figure}[!ht]
\includegraphics{../figs/mean-ar-by-regime-change-type-placebo.pdf}
\caption{Time-shifted placebo sensitivity analysis of mean event day abnormal return by type of regime change}
\label{fig:mean-event-day-ar-by-regime-change}
\end{figure}

@@

\newpage
\pdfbookmark[1]{References}{References}
\bibliography{bibliography}

>>>>

\singlespacing
\begin{abstract}
\noindent
We examine the relationship between political instability and daily returns of national stock indices in every country that experienced a coup, assassination, or resignation for which daily data is available. A common hypothesis is that political instability discourages investment and reduces economic growth. By contrast, we find that while financial volatility increases dramatically following ``irregular'' regime changes, different types of regime change have disparate effects on stock returns. We use daily financial data and a constant mean return event study model to show that abnormal returns following resignations are large and positive (4\%), while abnormal returns following assassinations are negative and smaller in magnitude (2\%). The impact of coups tends to be negative (2\%), but some events result in positive abnormal returns of 10\% or more. We also find that volatility increases during times of protest preceding resignations, but that no clear directionality is present. We therefore find that the expected direction and magnitude of abnormal returns is dependent on the type of political event and its expected impact on economic policy.
\end{abstract}
\doublespacing

@@

In this study, we examine the impact of ``irregular'' regime changes and public protests on economic performance and remain agnostic about the direction of the effect. In contrast to the aforementioned cross country studies---but like \citet{girardi2018institution}---we conduct our analysis using daily financial data. Market expectations in the form of stock market returns are used as an indicator of whether the market views different types of potentially destabilizing political events as ``good'' for economic growth. The advantage of this event study approach is that it allows us to use time-series variation in stock returns to estimate treatment effects and mitigate the endogeneity problems that are nearly ubiquitous in cross country regressions.\footnote{A growing body of work uses event studies to assess political phenomenon.  For example, political events have been used to estimate the effect of political connections on firm value \citep[e.g.][]{fisman2001estimating,faccio2006politically,goldman2009politically}. Studies in ``forensic economics" have used abnormal returns from political events to locate or examine transactions such as insider trading \citep{dube2011coups}, illegal arms trading \citep{dellavigna2010detecting}, and the impact of hostilities on the financial performance of diamond mining firms in Angola \citep{guidolin2007diamonds}.} We analyze 13 coups, 8 assassinations, 15 forced resignations, 11 public protests, and 1 natural experiment provided by the failed 2002 coup in Venezuela. Our study is therefore not only able to use an event-study approach to enhance the reliability of the estimates, but is not limited to a single case study as in \citet{girardi2018institution}. The relatively large sample enhances external validity and allows us to show that different types of political events have disparate economic effects.

We find that coups, assassinations, resignations, and public protests cause large increases in financial volatility. Abnormal returns following resignations are large and positive (4\%) while abnormal returns following assassinations are negative and smaller in magnitude (2\%). The average impact of coups is negative (2\%); however, stock prices increased dramatically during the failed overthrow of Hugo Chavez's socialist regime and decreased by approximately the same magnitude following his return to leadership. The effect of public protests on mean stock returns is uncertain, with some protests such as those during the Egyptian Revolution depressing them while others have the opposite effect.

@@

The failed coup attempt against Hugo Chavez provides a natural experiment that allows us to examine the effects of both pro-business and anti-business regime changes separately from simple uncertainty because investors reacted to an expected regime change twice: first, when Chavez was ousted, and second, when he was reinstated. On the evening of April 11, 2002, coup plotters removed Chavez from office and later detained him. Pedro Carmona, a Venezuelan economist and business leader, was named the transitional President of Venezuela. Two days later, on April 13, 2002, a popular uprising led to Chavez's reinstatement as president. This provides an estimate of the market's valuation of a transition from the Chavez regime to the Carmona regime and its valuation of a transition from the Carmona regime back to the Chavez regime. By extension, it provides an estimate of the impact of a shift from a left-wing populist government to a pro business regime in an emerging market.

\autoref{fig:AR-Ven} provides graphical evidence on the effect of the coup attempt. The top panel shows CARs for the 10 days prior to and following the event, along with 95\% confidence intervals. The daily ARs and corresponding confidence intervals are displayed in the bottom panel. The abnormal return on April 12, the first trading day in which investors could react to the coup, was 10\%. The market reacted in the opposite direction to Chavez's reinstatement as president: the abnormal return on the next trading day, April 15, 2002, was -8\%.

The results in the figure are particularly striking given the discrepancy between the ARs on event days 0 and 1 and all other days. Consistent with our earlier findings that coups tend to have pre and post-event CARs that are statistically indistinguishable from zero, the only days on which the 2002 Venezuelan coup d'etat attempt ARs are statistically different from zero is on event days 0 and 1 after the coup attempt. The almost 0\% 10-day CAR preceding the coup makes this an ideal case as it implies that investors were unaware of the coup plot. The unexpected nature of the events implies that both regime changes occurred in an as-if random fashion from the perspective of investors, increasing our confidence that the abnormal returns capture the true effect of the Chavez to Carmona and Carmona to Chavez regime changes on stock returns.

@@

The results in \autoref{tab:AR-ass} are produced from analyses identical to those in table \autoref{tab:AR-coups} but for successful assassinations rather than coups. Like the majority of coups, there is evidence that assassinations decrease stock prices. The mean event day abnormal return is negative and statistically different than zero. However, the result is driven by five events: the shooting of U.S. President William McKinley on September 6, 1901; the assassination of U.S. President John F. Kennedy on November 22, 1963; the assassination of Indian Prime Minister Indira Gandhi on October 31, 1984; the suicide bombing that killed Sri Lankan president Ranasinghe Premadasa on May 1, 1993; and the assassination of Israeli Prime Minister Yitzhak Rabin on the evening of November 4, 1995.

@@

These results are consistent with our hypothesis that the nature of the political event and its expected impact on policy matters. While the mean effect of assassinations is negative, it is smaller in magnitude than for coups. Unlike a coup, an assassination may not necessarily be expected to cause immediate change in economic policy, particularly in the presence of an institutionalized line of succession. As such, we would expect CARs to be negative due to increased instability and uncertainty, but smaller in magnitude to a coup or resignation due to greater expectations of policy inertia.

@@

In contrast to coups and assassinations, abnormal returns following resignations are large and positive (see \autoref{tab:AR-resignations}). The mean event day abnormal return is over 4\% and the positive returns are persistent and grow larger over time (mean 20-day CAR $\approx$ 12\%). Furthermore, event day ARs are only negative and statistically significant at even the ten percent level in two out of the fifteen resignations (Pakistan on April 19, 1993 and Tunisia on June 31, 2011).

@@

For example, consider Ferdinand Marcos' resignation from office as President of the Philippines in February 1986. Prior to his resignation, the Philippine regime was known for rampant corruption, crony capitalism, extreme inequality, high unemployment, failed import substitution industrialization policy, and oligarchic control of the economy \citep{overholt1986rise, traywick2014}. In fact, the Philippines was the least preferred site for foreign investment amongst East Asian capitalist economies and possessed one of the worst capital investment to economic output ratios in Asia \citep{overholt1986rise}. Marcos held a snap presidential election on February 7, 1986, in which he declared victory despite overwhelming evidence of electoral fraud. Public protests ensued, and two weeks later the military withdrew its support of the Marcos regime \citep{lee2009armed}. Marcos was replaced by his electoral opponent, Corazon Aquino, who had run on a platform of economic liberalization and elimination of crony capitalism \citep{villegas1987philippines}. This event was associated with an approximately 13\% positive event day AR.

@@

The short-term impact of the Egyptian Revolution on the economy was disastrous. As shown in \autoref{fig:CAR-Egypt}, abnormal returns on the Egyptian Stock Exchange Index (EGX 30) were around -7\% on January 26th and -10\% the day after. To prevent further decline during the uprising, the Egyptian Stock Exchange closed at the end of trading on January 27th. President Mubarak resigned on February 11, but the market remained closed until March 23, when CARs declined by another 9\%, before rebounding slightly thereafter.

@@

\begin{figure}[!htb]
\includegraphics{../figs/mean-volatility.pdf}
\caption{Mean of volatility estimates from GARCH(1,1) models}
\label{fig:volatility}
\end{figure}

@@

\begin{figure}[!htb]
\centering
\begin{subfigure}{\textwidth}
\centering
\includegraphics[max size={.8\textwidth}{.8\textheight}]{../figs/mean-car-by-regime-change-type.pdf}
\caption{Actual event date} \label{subfig:mean-car-by-regime-change-actual}
\end{subfigure}
\begin{subfigure}{\textwidth}
\centering
\includegraphics[max size={.8\textwidth}{.8\textheight}]{../figs/mean-car-by-regime-change-type-placebo.pdf}
\caption{Event date shifted forward by one year} \label{subfig:mean-car-by-regime-change-placebo}
\end{subfigure}
\caption{Mean cumulative abnormal returns by type of regime change}\label{fig:mean-car-by-regime-change}
\end{figure}

To ensure that results of the placebo test based on 1-year did not occur by chance, \autoref{fig:mean-event-day-ar-by-regime-change} plots event-day abnormal against the number of days shifted. There are a few instances in which there are statistically significant (at the 5\% level) ARs in the same direction as those on the actual event day, but they are always smaller in magnitude than the ARs estimated using the actual event date. Most of the statistically significant placebo estimates also occur when dates are shifted within the post-event window (days 0-20), a period during which stock returns remain volatile and, in the case of resignations, there is a consistent upward trend in the CARs. When dates are shifted forward further ($\geq$ 110 days), the event day AR is only statistically significant in one case (day 365 for assassinations). Overall, these results reinforce the main results: the ARs on the actual event date capture most of effect of the regime change, although effects can sometimes persist in the short event window following the event date.

These figures also provide evidence on the extent to which regime changes appear to be unexpected. For instance, the CARs prior to assassinations and coups presented in \autoref{subfig:mean-car-by-regime-change-actual} tend to be close to zero, suggesting that investors were unaware that a negative event was likely to occur in the coming days. For resignations, CARs trend downward in the 10 days prior to the event; however, if investors were aware that a resignation were about to occur one would expect the pre-event CARs to be positive given the positive CARs observed in the post-event window. There is thus little evidence that the CARs in the post-event window are not capturing most of the effect caused by the regime changes.

@@

\begin{table}[!ht]
\caption{Non-parametric tests of the impact of regime changes} \label{tab:non-parametric}
\vspace{-5pt}
\footnotesize
\begin{center}
\begin{threeparttable}
\begin{tabular*}{\textwidth}{l@{\extracolsep{\fill}}d{1.3}d{1.3}d{1.3}d{1.3}d{1.3}d{1.3}d{1.3}}
  \hline
  \hline
&\multicolumn{3}{c}{Regime Change Country}&\multicolumn{3}{c}{Synthetic Control Portfolio}&\multicolumn{1}{c}{\multirow{2}{*}{Wilcoxon}}\\
\cmidrule(r){2-4} \cmidrule(r){5-7}
&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Rank}&\multicolumn{1}{c}{Sign}
&\multicolumn{1}{c}{Mean}&\multicolumn{1}{c}{Rank}&\multicolumn{1}{c}{Sign}
&\multicolumn{1}{c}{Rank Test}\\
\multicolumn{1}{c}{Event Type}&\multicolumn{1}{c}{CAR (0,0)}&\multicolumn{1}{c}{p-value}&\multicolumn{1}{c}{p-value}
&\multicolumn{1}{c}{CAR (0,0)}&\multicolumn{1}{c}{p-value}&\multicolumn{1}{c}{p-value}
&\multicolumn{1}{c}{p-Value}\\
  \hline
\ExpandableInput{../tables/non-parametric-ar0.txt}
   \hline
   \hline
\end{tabular*}
\scriptsize
Notes: Estimates for assassinations do not include the assassination of U.S. president William McKinley in 1901 because no control portfolios are available.
\end{threeparttable}
\end{center}
\end{table}

As shown in \autoref{tab:non-parametric}, the mean event day abnormal returns for coups, assassinations and resignations are all statistically different from zero at the 1\% level using the rank test statistic and the abnormal returns for coups and assassinations are significant at at least the 10\% level using the sign test. In addition, abnormal absolute returns for all events are statistically significant at at least the 5\% level using both the rank and sign tests. On the other hand, the event day abnormal returns for the control portfolio are never statistically different from zero at even the 10\% level using the rank or sign tests. Finally, the difference in means between the regime change country and the control portfolio are statistically different from zero for coups (1\% level), assassinations (10\% level), resignations (5\% level), and all events combined (1\% level) when using two-sided p-values from the Wilcoxon rank test.\footnote{The Wilcoxon rank test is a non-parametric statistical technique that can be used to compare differences between matched samples.} In sum, the synthetic control and small sample tests suggest that the main results are not a result of deviations from normality or confounding world events.

\section{Discussion and Conclusion}
Max Weber (\citeyear{weber1947theory}) developed the concept of ``charismatic leadership'' whereby a leader's authority stems from their charisma. Such rule is often found in autocracies with weak institutions and few constraints placed on the power of leaders. \citet{jones2005leaders} show that leaders can have particularly large impacts on economic growth in these settings. It is therefore not surprising that there is often a strong temptation to overthrow ``bad" leaders, even by undemocratic means.

@@

The results are consistent with the idea that perceptions of government competence and changes in government have large impacts on investor confidence. But although the effect of regime changes and protests on stock volatility is substantial, the effect on the direction of stock returns is less certain. Abnormal returns following resignations are large and positive (4\%), abnormal returns following assassinations are negative and smaller in magnitude (2\%), and abnormal returns following coups also tend to be negative (2\%). However, the returns surrounding the failed 2002 coup that temporarily ousted Hugo Chavez from power in Venezuela show that coups can sometimes have large positive economic effects, which is consistent with estimates from \citet{girardi2018institution} showing  the election of Allende in Chile and the subsequent coup replacing him increased stock market valuations.

In total, the evidence consequently suggests that abnormal returns tend to be positive following resignations because the regime change likely replaced ``bad'' leaders, and coups can result in positive returns in settings where socialist governments are replaced with more pro-business ones. Conversely, autocratic coups may increase fear of capital appropriation or the implementation of poor economic policies, while assassinations may be merely regarded as highly uncertain events.

As in any study, there are a number potential avenues for future research. First, more research is needed to identify the pro- and anti-market characteristics of regime changes. There may be features of regime changes that are perceived as pro- or anti-market that are more nuanced than those identified in this paper. A greater understanding of these features would allow researchers to better predict the effects of future regime changes on stock returns. Additionally, it would be useful to identify the mechanisms through which regime changes effect stock returns. In the case of Chile,  \citet{girardi2018institution} show that increasing stock market valuations were caused by changes in private property rights rather than economic growth prospects or wage costs. Research focusing on the mechanisms in other settings would help generalize these findings. Lastly, it would be helpful to determine the extent to which stock market returns translate to broader economic development outcomes. \citet{meyersson2016political} has made progress on this front by examining the impact of coups on a number of outcomes in addition to economic growth including investment, debt, inflation, infant mortality, and years of schooling. It would be fruitful to examine whether the direction of the effects of different types of regime changes on these outcomes are consistent with their stock market effects, or if investor perceptions are at odds with certain development goals. Such research can help enlarge the body of evidence on the extent to which regime changes cause institutional and political change, and, in turn, have significant consequences for economic development.

@@

\pagebreak

\subsection{Time-shifted placebo test}

\begin{figure}[!htb]
\includegraphics{../figs/mean-ar-by-regime-change-type-placebo.pdf}
\caption{Time-shifted placebo sensitivity analysis of mean event day abnormal return by type of regime change}
\label{fig:mean-event-day-ar-by-regime-change}
\end{figure}

\pagebreak

@@

\clearpage
\pagebreak

\pdfbookmark[1]{References}{References}
\bibliography{bibliography}
